\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{titlesec}
\usepackage{fancyhdr}
\usepackage{parskip}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{tcolorbox}

\hypersetup{colorlinks=true, linkcolor=blue!60!black, urlcolor=blue!60!black}
\setlist[itemize]{topsep=4pt,itemsep=3pt,leftmargin=1.4em}
\setlist[enumerate]{topsep=4pt,itemsep=3pt,leftmargin=1.6em}
\titleformat{\section}{\large\bfseries}{\thesection}{0.6em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.6em}{}

\lstdefinestyle{bashstyle}{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{black!3},
  frame=single,
  rulecolor=\color{black!15},
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  breaklines=true
}

\pagestyle{fancy}
\fancyhf{}
\lhead{Project \#4: Find and Fix Bugs with Fuzzing}
\rhead{Secure System Engineering and Management}
\cfoot{\thepage}

\begin{document}

\begin{center}
    {\LARGE \textbf{Project \#4: Find and Fix Bugs with Fuzzing}}\\[0.4em]
    {\large Secure System Engineering and Management}\\[0.4em]
    Due date: \textit{XXXXXX}
\end{center}

\section*{Overview}
In this project, you will use coverage-guided fuzzing to discover, reproduce, diagnose, and fix memory-safety bugs in a buggy PNG-processing target. The target is based on \texttt{libpng} and is packaged in a containerized local OSS-Fuzz workflow so that everyone uses the same build and runtime environment.

The goal is to practice the core workflow used in modern software security engineering:
\begin{itemize}
    \item run a fuzzer against a target,
    \item interpret sanitizer crash reports,
    \item trace the bug to its root cause in source code,
    \item patch the bug correctly, and
    \item verify that the crashing input no longer crashes without breaking valid behavior.
\end{itemize}

This project emphasizes debugging and repair, not just crash discovery. You are expected to read C/C++ source, understand AddressSanitizer reports, and implement correct fixes.

\section*{What You Are Given}
You are provided with a repository containing:
\begin{itemize}
    \item a buggy course-owned fuzz target based on \texttt{libpng},
    \item an OSS-Fuzz project configuration for local Docker-based fuzzing,
    \item a small PNG seed corpus to help the fuzzer reach valid parsing paths,
    \item a fuzz harness and intentionally seeded memory-safety bugs, and
    \item helper scripts for building, fuzzing, reproducing crashes, and post-fix regression testing.
\end{itemize}

You should not replace the provided target with a different target. Your work should focus on understanding, fixing, and validating the bugs in the provided code.

\section*{Learning Objectives}
By completing this project, you will:
\begin{itemize}
    \item run a containerized fuzzing workflow with a real parser target,
    \item interpret AddressSanitizer reports for memory-safety bugs,
    \item reproduce crashes from specific inputs,
    \item diagnose and patch use-after-free and double-free bugs,
    \item understand how bug discovery can be sequential, and
    \item validate that fixes eliminate the crash while preserving correct behavior on valid inputs.
\end{itemize}

\section*{Setup Instructions (Read Carefully)}
You must use the provided Docker/OSS-Fuzz workflow. Native local builds outside the containerized environment are \textbf{not supported}.

\subsection*{Step 1: Get the Code}
Clone the course repository from the link provided on Canvas, then enter the repository root:

\begin{lstlisting}[style=bashstyle]
git clone <repo-url>
cd ssem-project4-fuzzing
\end{lstlisting}

The repository root should contain at least:
\begin{itemize}
    \item \texttt{libpng-course/}
    \item \texttt{oss-fuzz/}
    \item \texttt{seed\_corpus/}
    \item \texttt{scripts/}
\end{itemize}

\subsection*{Step 2: Install Prerequisites}
You need:
\begin{itemize}
    \item Git
    \item Docker Desktop (Mac/Windows) or Docker Engine (Linux)
    \item Python 3
\end{itemize}

Confirm the following work in a terminal:
\begin{lstlisting}[style=bashstyle]
git --version
docker --version
python3 --version
\end{lstlisting}

\subsection*{Step 3: Build the Project}
From the repository root, run:
\begin{lstlisting}[style=bashstyle]
./scripts/build.sh
\end{lstlisting}

This script:
\begin{itemize}
    \item builds the local OSS-Fuzz project named \texttt{libpng-course},
    \item builds the fuzz target, and
    \item installs the provided PNG seed corpus in the correct location.
\end{itemize}

If the build script asks whether to use a saved image or cached base images, always select \textbf{N}. Use the fresh rebuild path.

The helper scripts explicitly use \texttt{python3}; if your system also has a \texttt{python} command, ignore it and use the provided scripts as written.

\subsection*{Step 4: Run the Fuzzer}
From the repository root, run:
\begin{lstlisting}[style=bashstyle]
./scripts/fuzz.sh 30
\end{lstlisting}

This runs the fuzzer for 30 seconds. You may increase the time if needed. With the provided seed corpus, the fuzzer should usually expose the first required bug quickly and print an AddressSanitizer report.

\subsection*{Step 5: Reproduce an Input}
To reproduce a crash on a specific input, run:
\begin{lstlisting}[style=bashstyle]
./scripts/reproduce.sh seed_corpus/pngnow.png
\end{lstlisting}

For some bugs, one of the provided seed inputs may already act as a reproducer. For generated crash inputs, replace the argument with the appropriate file path.

\subsection*{Step 6: Post-Fix Regression Check}
After fixing the required bugs, run:
\begin{lstlisting}[style=bashstyle]
./scripts/test_valid.sh
\end{lstlisting}

This verifies that the provided valid PNG inputs still execute successfully after your fixes.

\subsection*{Debugging Common Issues}
\begin{itemize}
    \item \textbf{Docker not running:} Start Docker Desktop or the Docker daemon, then rerun the commands.
    \item \textbf{Permission denied on scripts:} Run \texttt{chmod +x scripts/*.sh}.
    \item \textbf{Seed corpus not used:} Re-run \texttt{./scripts/build.sh}. The build script installs the seed corpus automatically.
    \item \textbf{Build succeeds but fuzzing finds nothing:} Confirm that \texttt{fuzz.sh} prints \texttt{Using seed corpus: course\_png\_fuzzer\_seed\_corpus.zip}.
    \item \textbf{Docker build fails with ``not enough free space'' or apt errors:} Your Docker environment may be out of disk space. Try \texttt{docker system prune -a -f} and \texttt{docker builder prune -a -f}, then rerun \texttt{./scripts/build.sh}.
\end{itemize}

\section*{Project Instructions}
This is a debugging and code-repair assignment.

You must use fuzzing to identify and fix \textbf{two distinct seeded bugs} in the provided target. These bugs are designed to be discovered \textbf{sequentially}:
\begin{enumerate}
    \item Run the fuzzer and identify the first bug.
    \item Reproduce the crash and patch the bug.
    \item Rebuild the project.
    \item Run the fuzzer again to expose the next bug.
    \item Reproduce and patch the second bug.
\end{enumerate}

A complete solution requires more than simply running the fuzzer: you must understand each crash, locate the buggy code, implement a correct fix, and verify that your fix preserves expected behavior.

You should implement your fixes in:
\begin{lstlisting}[style=bashstyle]
libpng-course/course/course_vuln.c
\end{lstlisting}

Do not replace the provided target with a different target, and do not substantially rewrite the provided fuzzing infrastructure.

\subsection*{What Counts as a Correct Fix?}
A correct fix:
\begin{itemize}
    \item removes the memory-safety error,
    \item modifies the buggy target logic rather than simply deleting large parts of the code,
    \item does not reject all inputs indiscriminately, and
    \item preserves normal behavior on the provided valid PNG inputs.
\end{itemize}

\section*{Required Writeup}
For each bug you fix, answer the following:
\begin{enumerate}
    \item What input triggered the crash?
    \item What sanitizer error was reported?
    \item Which function and line caused the bug?
    \item What is the root cause?
    \item How did you fix it?
    \item How did you verify the fix?
\end{enumerate}

Then write a short reflection addressing both of the following:
\begin{enumerate}
    \item What did fuzzing help you discover that would have been harder to find by inspection alone?
    \item What are the limitations of coverage-guided fuzzing in this project, especially for reaching deeper semantic conditions?
\end{enumerate}

\section*{Submission Instructions}
Submit the following on Canvas:
\begin{itemize}
    \item your patched \texttt{libpng-course/course/course\_vuln.c} file \textbf{or} a unified patch/diff against the original file,
    \item your writeup (PDF or Markdown), and
    \item a small folder named \texttt{reproducers/} containing two files, one reproducing input for each of the two bugs you fixed.
\end{itemize}

Name the reproducer files clearly, for example:
\begin{lstlisting}[style=bashstyle]
reproducers/bug1_input
reproducers/bug2_input
\end{lstlisting}

These reproducer files are raw binary inputs. They are \textbf{not} screenshots or pasted text. They should be the exact files you use with \texttt{./scripts/reproduce.sh}. After your fixes are applied, these same reproducer files should no longer trigger a sanitizer failure.

Do not submit Docker images, corpora generated during fuzzing, or unrelated build artifacts.

\section*{Grading}
This project is graded using a combination of automated checks and manual review.

\begin{center}
\begin{tabular}{>{\raggedright\arraybackslash}p{0.62\linewidth} >{\raggedleft\arraybackslash}p{0.18\linewidth}}
\toprule
Component & Weight \\
\midrule
Setup works; code builds and runs in the provided workflow & 15\% \\
Bug 1: reproducer, diagnosis, and fix & 25\% \\
Bug 2: reproducer, diagnosis, and fix & 25\% \\
Validation that crashing inputs no longer crash and valid PNGs still work & 20\% \\
Clarity and correctness of the written explanations & 15\% \\
\bottomrule
\end{tabular}
\end{center}

Correct code is necessary but not sufficient. Clear reasoning and validation matter.

\section*{Academic Integrity}
You may discuss high-level ideas with classmates, but all code, debugging, and writing must be your own. Do not share completed solutions, crash inputs with explanations, or patched source files.

\section*{Use of Generative AI Tools}
Students are permitted to use generative AI tools (e.g., large language models) to assist with aspects of this project, including:
\begin{itemize}
    \item understanding sanitizer reports, compiler errors, or unfamiliar C/C++ syntax,
    \item debugging code or exploring alternative bug fixes,
    \item clarifying how fuzzing, memory errors, or Docker-based workflows operate,
    \item brainstorming approaches to reproducing and validating crashes, and
    \item improving the clarity of written explanations.
\end{itemize}

However, all submitted work must reflect your own understanding and reasoning. In particular:
\begin{itemize}
    \item You may not submit code, patches, or explanations that you do not understand.
    \item You may not outsource the core debugging, repair, or interpretation of the bugs to an AI system.
    \item You should be able to explain what each bug was, why your fix is correct, and how you verified it.
\end{itemize}

You are responsible for the correctness of any code, reproducer inputs, and written explanations you submit, regardless of whether an AI tool was used to help generate them. We may ask questions about your fixes, reproducer inputs, or debugging process to assess understanding.

This project emphasizes debugging, reasoning, validation, and explanation. Over-reliance on generative tools without understanding may negatively affect grading, especially in categories related to diagnosis, correctness, and written explanation.

\section*{Notes on Fuzzing Engine Choice}
This project uses \texttt{libFuzzer} in the provided workflow because it integrates smoothly with the local OSS-Fuzz setup and is easier to run consistently across different laptop environments through Docker. AFL++ is an important modern standalone fuzzer and may be discussed in class, but all required grading for this project uses the provided \texttt{libFuzzer}-based workflow.

\end{document}
